---
title: "Go-No Go Tutorial"
author: "Milla Pihlajamaki"
date: "27 5 2022"
output: pdf_document
---

# Go-No Go Task Tutorial

This tutorial will demonstrate how the *gonogo* package is used. The package contains two functions: play_gonogo() for playing the Go-No Go Task, and check_rt() for checking for irregularities in the output data, specifically the reaction time column. 

## Go-No Go Task in Short

The Go-No Go Task is a widely used test to measure inhibitory control, a cognitive process that enables humans to cancel motor activity after its initiation. It requires the participant to perform an action given certain stimuli (Go stimuli), and inhibit that action under a different set of stimuli (No Go stimuli).

There are two parameters in the experimental design that are especially important: the length of each trial and the relative proportion of the Go and No-Go trials. Fortunately, both these parameters can be easily manipulated in the gonogo() function: length of the trial with the *inter* argument, and the relative proportion of the Go and No-Go trials with the *prb* argument.

In addition to these two arguments, you can specify the participant *id* (name or unique id number), *n_trial* (number of trials), *n_block* (number of blocks), and *stimuli* (the Go and No Go stimuli). 

## How to Use the play_gonogo() Function

The following code gives an example of how the play_gonogo() function can be used.

```{r Install the Package, message=FALSE, warning=FALSE}

# Load package
# devtools::install_github("Programming-The-Next-Step-2022/GoNoGoTask/gonogo")
library(gonogo)

```


```{r The play_gonogo() function, eval=FALSE}

# Run the Go-No Go Task and save the output in an object
p1_data <- play_gonogo(id = "p1",  # id
                       n_trial = 5,  # number of trials
                       n_block = 3,  # number of blocks
                       stimuli = c("A", "X"),  # Go and No Go stimuli
                       inter = 0.6,  # interval (length of trials)
                       prb = c(0.8, 0.2))  # relative proportion

```


## Example Screens from the Task

Below you can find examples of how the task looks: the welcome screen, information about the number of trials and blocks, and finally an example of how the stimuli are presented during the task.

```{r Example Screens from the Task, echo=FALSE, height = 0.8}

# Welcome screen
id <- "p1"
draw_screen(paste0("Welcome to the Go-No Go Task ", id,
                   "!\nPress Enter to continue."), cex = 1.3)

# Information about number of trials and blocks
n_block <- 3
n_trial <- 5
draw_screen(paste0("The task consists of ", n_block, " blocks, each consisting of ", n_trial, " trials.\nThere is is a break between each pair of blocks.\nBefore starting the task, there is one practice block of 10 trials.\n Press Enter to start the practice trials."),
             cex = 0.75)

# Example of how the task looks like
draw_screen("A", 3)


```



```{r Create dataset, include=FALSE}

p1_data <- data.frame(id = rep("p1", 15),
                      response = c("none", "none", "space", "space", "none", "space", "none", "space", "space", "none", "none", "space", "none", "space", "none"),
                      correct = c(1,1,1,1,0,0,0,1,1,1,0,1,0,1,0),
                      SDT = c("correctrejection", "correctrejection", "hit", "hit", "miss", "falsealarm", "miss", "hit", "hit", "correctrejection", "miss", "hit", "miss", "hit", "miss"),
                      rt = c(NA, NA, 0.47666382, 0.55146193, NA, 0.56809711, NA, -0.01646210, 0.59251689, NA, NA, -0.01505090, NA, -0.02203418, NA),
                      stimulus = c("X", "X", "A", "A", "A", "X", "A", "A", "A", "X", "A", "A", "A", "A", "A"),
                      block = rep(c(1, 2, 3), each = 5))

```


## What the Output Data Looks Like

```{r Show data, echo=FALSE}

print(p1_data)

```


## How to Read the Output

The play_gonogo() function returns a dataframe consisting of n_trial*n_block (number of trials times number of blocks) rows and seven columns:


**id** = participant's name or id as specified 

**response** = response key used on the trial (space when participant responded, none when no response was given)

*correct* = whether the response was correct or not (1=correct, 0=incorrect)

**SDT** = responses categorized according to Signal Detection Theory; [click here to read more about Signal Detection Theory](https://en.wikipedia.org/wiki/Detection_theory)

**rt** = reaction time in seconds (NA when participant did not respond during that trial)

**stimulus** = the stimulus shown on the trial

**block** = the block number


## The check_rt() Function

It appears that there is something weird going on in the reaction time column:

```{r Print Reaction Times, echo=FALSE, message=FALSE, warning=FALSE}

hist(p1_data$rt, breaks = 25, col = "lightblue", title = "Reaction times (in seconds)", xlab = NULL)

```

There are some observations that are very close to (or below) 0. This is the case when the participant takes too long to respond to a stimulus, and their response "leaks" into the next trial. As the function corrects for the time that it takes to run the function to gain more accurate reaction times by subtracting a constant from the reaction times, this might sometimes lead to very small or negative values.

Fortunately, the gonogo package comes with a function that helps spot irregularities in reaction times: check_rt()! The check_rt() function points our unsually long distances between sorted observations in your reaction time data. It has two arguments, *data* (specify your dataframe), and *ratio* (the ratio of the longest distance between sorted data points and the range of the data). See below:

```{r How to Use the check_rt() function}

check_rt(data = p1_data, ratio = 1/3)

```

## End of the Tutorial!
